---
title: "Midterm Exam"
subtitle: "Fall 2023"
author: "Prof. Bell"
# format:
#   html:
#     embed-resources: true
#     # code-tools: true
format: gfm
execute:
  include: false
  warning: false
  message: false
editor_options: 
  chunk_output_type: inline
---

The Bipartisan Infrastructure Law of 2021 is a signature achievement of the Biden Administration. This act includes \$550 billion in new spending on infrastructure improvements, such as repairing bridges that are becoming unsafe for travel.

There are more than 600,000 bridges in the United States. Using data from the Federal Highway Authority, we will answer the questions: **Which bridges should the Biden Administration prioritize with new infrastructure spending?**

Please use the "Mid-term Exam" project on Posit Cloud for this work.

You must submit your exam as a rendered Quarto document (HTML preferred, PDF and Word acceptable). Please ensure that all code used to generate the document (including your `setup` chunk) are visible in the report. However, extraneous code that is not required for answering the questions but that appears in your Quarto document will result in a deduction (for example, `glimpse()` should not appear anywhere in your rendered report).

Please turn in your `.qmd` file, the `_files` directory, and your rendered Quarto document as a `.zip` folder on Blackboard. When you export multiple items from Posit Cloud at once, it will automatically be exported as a `.zip` file.

```{r}
#| label: setup
library(tidyverse)
```

1. Load `bridges.csv` and `fips.csv` from the `data` folder. (5 points)

```{r}
#| label: question1

bridges <- read_csv("data/bridges.csv", name_repair = "universal")
fips <- read_csv("data/fips.csv")
```

2. Create a table showing the number of bridges that are part of the National Highway System vs. not. (5 points)

```{r}
#| label: question2

table(bridges$National.Highway.System)
```

3. Let's wrangle this data. Create a new column indicating the current age of each bridge, and keep only bridges that have traffic of at least 25,000 vehicles per year. (5 points)

```{r}
#| label: question3

bridges <-
  bridges |>
  mutate(age = 2023 - Year.Built) |>
  filter(Traffic >= 25000)
```

4. Create a nicely-formatted graph showing the age distribution of bridges in the United States. (10 points)

```{r}
#| label: question4

ggplot(bridges) +
  geom_histogram(aes(x = age))
```

5. Bridges are rated as one of three conditions: Good, Fair, or Poor. Create a nicely-formatted graph that shows the ages of bridges in each condition. Is there a relationship between age and condition? (10 points)

```{r}
#| label: question5

ggplot(bridges) +
  geom_boxplot(aes(x = Condition, y = age))

# Bridges in worse condition tend to be older.
``` 

6. Join the `fips.csv` data to the bridges data, keeping all of the rows in the bridges data. (5 points)

```{r}
#| label: question 6

joined <- left_join(bridges, fips, by = join_by(FIPS == fips))
```

7. Which five states have the greatest percent of bridges in "Poor" condition? (10 points)

    *Hint: Remember that the mean of 1s and 0s (or TRUEs and FALSEs) is a percent.*

```{r}
#| label: question7

joined |>
  group_by(state_name) |>
  summarize(poor_pct = mean(Condition == "Poor", na.rm = TRUE)) |>
  arrange(-poor_pct) |>
  filter(!is.na(state_name)) |>
  slice_head(n = 5)
```
