---
output: 
    stevetemplates::statement:
    latex_engine: pdflatex
geometry: margin=1in

title: "Assignment #2"
author: Prof. Bell
email: Spring 2023

fontfamily: cochineal
fontsize: 11pt
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = F, include = F)
```

Install the `moderndive` package and load the `house_prices` data. For this homework, please produce a nicely-formatted graph that helps answer each of the following questions. Put your graphs in a Word document (or other word processing software) and **include a brief (1-2 sentence) answer to the question.**

```{r, message = F}
# Not run
# install.packages("moderndive")
library(tidyverse)
library(moderndive)
data(house_prices)
```

1. Create a new variable for the month that a home was sold using this command: `house_prices$month <- lubridate::month(house_prices$date, label = T)`. What is the most popular month for home sales?

```{r}
house_prices$month <- lubridate::month(house_prices$date, label = T)

ggplot(data = house_prices) +
	geom_bar(aes(x = month))
```

2. In 1900, the population of King County was 110,053; in 2020, it was 2.3 million. Using home construction as an indicator of population growth, in what decade did the population of King County start to grow dramatically?

```{r}
ggplot(data = house_prices) +
	geom_histogram(aes(x = yr_built), color = "white", binwidth = 10)
```

3. Do homes built after 1990 tend to have more bathrooms per bedroom than homes built before 1990? (Hint: You will need to use a logical statement when constructing your graph.)

```{r}
ggplot(data = house_prices, aes(x = bedrooms, y = bathrooms, color = yr_built > 1990)) +
	geom_point(alpha = .5) +
	geom_smooth(method = "lm")
```

4. Of the ten zip codes with the most homes sold in 2015, which was the most expensive zip code for new homes? (Please report prices in $1000s, so divide price by 1000.)

The zip codes are:
```{r, include = T}
dplyr::count(house_prices, zipcode) %>%
	dplyr::arrange(desc(n)) %>%
	dplyr::slice_head(n = 10) %>% 
	select(zipcode)
```

```{r}
top_zips <- house_prices[house_prices$zipcode %in% c(98103, 98038, 98115, 98052, 98117, 98042, 98034, 98118, 98023, 98006),]

ggplot(data = top_zips) +
	geom_boxplot(aes(x = factor(zipcode), y = price/1000))
```

5. What is the relationship between the age of a home and it's price? Is age more or less important than condition in determining the price? (Please report prices in $1000s, so divide price by 1000.)

```{r}
ggplot(data = house_prices, aes(x = 2015-yr_built, y = price/1000)) +
	geom_point(alpha = .2) +
	geom_smooth(method = "lm") +
	facet_wrap(~ condition)
```
