---
output: 
    stevetemplates::statement:
    latex_engine: pdflatex
geometry: margin=1in
graphics: yes

title: "Assignment #2"
author: Prof. Bell
email: Fall 2023

fontfamily: cochineal
fontsize: 11pt
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = F, include = F, fig.height = 4, fig.width = 6, fig.align = "center")
```

For this homework, please produce a nicely-formatted graph that answers each of the following questions. A nicely-formatted graph includes:

* Proper axis and legend labels
* A title, preferably centered
* A caption indicating the source of the data
* Any other aesthetic elements that make the graph visually pleasing

**Include a brief (1-2 sentence) written answer to the question as well.**

You must submit your homework as a rendered Quarto document (HTML preferred, PDF and Word acceptable). Please ensure that all code used to generate the document (including your `setup` chunk) are visible in the report. However, extraneous code that is not required for answering the questions but that appears in your Quarto document will result in a deduction (for example, `glimpse()` should not appear anywhere in your rendered report).

Please turn in your `.qmd` file, the `_files` directory, and your rendered Quarto document as a `.zip` folder on Blackboard. When you export multiple items from Posit Cloud at once, it will automatically be exported as a `.zip` (If you cannot get the report to render, you may turn in only the `.qmd` file, but will lose points on the Quarto elements of the grading rubric.)

In your `setup` chunk, install the `moderndive` package and load the `house_prices` data. 

```{r, message = F}
# Not run
# install.packages("moderndive")
library(tidyverse)
library(moderndive)
data(house_prices)
```

1. Create a new variable for the month that a home was sold using this command:

    `house_prices$month <- lubridate::month(house_prices$date, label = T)`

    What is the most popular month for home sales?

```{r}
house_prices$month <- lubridate::month(house_prices$date, label = T)

ggplot(data = house_prices) +
	geom_bar(aes(x = month)) +
	labs(x = "Month",
			 y = "Number of Homes Sold",
			 title = "Number of Homes Sold Per Month",
			 caption = "Source: moderndive package") +
	theme_minimal() +
	theme(plot.title = element_text(hjust = .5))
```

2. In 1900, the population of King County was 110,053; in 2020, it was 2.3 million. In what decade did home construction in King County start to grow dramatically? (Hint: You do not need to create a new variable; explore the `binwidth` argument.)

```{r}
ggplot(data = house_prices) +
	geom_histogram(aes(x = yr_built), color = "white", binwidth = 10) +
	labs(x = "Decade",
			 y = "Number of Homes Sold",
			 title = "Number of Homes Sold Per Decade",
		   caption = "Source: moderndive package") +
	theme_minimal() +
	theme(plot.title = element_text(hjust = .5))
```

3. Design style has changed over the past century. Using a scatterplot, show how the number of bathrooms compares to the number of bedrooms. Do homes built on or after 1990 tend to have more bathrooms per bedroom than homes built before 1990?

    *1 bonus point: Add best fit lines to the scatterplot to show the strength of the relationship between bathrooms and bedrooms before and after 1990. Comment on these relationships.*

```{r}
ggplot(data = house_prices, aes(x = bedrooms, y = bathrooms, color = yr_built > 1990)) +
	geom_point(alpha = .5) +
	geom_smooth(method = "lm", se = FALSE) +
	labs(x = "Bedrooms",
			 y = "Bathrooms",
			 color = "Built after 1990?",
			 title = "Bedrooms and Bathrooms Pre- and Post-1990 Construction",
			 caption = "Source: moderndive package") +
	theme(plot.title = element_text(hjust = .5))
```

4. What is the relationship between the condition of a home and it's price? Please report prices in $1000s, so divide price by 1000. (For condition, 1 = Poor, 2 = Fair, 3 = Average, 4 = Good, and 5 = Very Good.)

    *2 bonus points: Compare the relationship between condition and price for older and newer homes (however you choose to define that) using facets. Comment on whether age impacts the relationship between condition and price.*

```{r}
ggplot(data = house_prices, aes(x = 2015-yr_built, y = price/1000)) +
	geom_point(alpha = .2) +
	geom_smooth(method = "lm") +
	facet_wrap(~ condition, nrow = 1) +
	labs(x = "Age of Home",
			 y = "Price ($1000s)",
			 title = "Price of Homes Sold by Age and Condition",
			 caption = "Source: moderndive package") +
	theme_bw(base_size = 8) +
	theme(plot.title = element_text(hjust = .5))
```
