---
title: "Assignment #4"
subtitle: "Fall 2023"
author: "Prof. Bell"
# format:
#   html:
#     code-tools: true
#     embed-resources: true
format: gfm
execute:
  warning: false
  message: false
editor_options: 
  chunk_output_type: inline
---

For this homework, we will use the `tidyverse` to estimate the favorability ratings of Joe Biden and Donald Trump among the U.S. public using data from Roper iPoll.

You must submit your homework as a rendered Quarto document (HTML preferred, PDF and Word acceptable). Please ensure that all code used to generate the document (including your `setup` chunk) are visible in the report. However, extraneous code that is not required for answering the questions but that appears in your Quarto document will result in a deduction (for example, `glimpse()` should not appear anywhere in your rendered report).

Please turn in your `.qmd` file, the `_files` directory, and your rendered Quarto document as a `.zip` folder on Blackboard. When you export multiple items from Posit Cloud at once, it will automatically be exported as a `.zip`.

```{r}
#| label: setup

library(tidyverse)
```

1. Download the data on the [August 2023 ABC News/Ipsos poll](https://doi.org/10.25940/ROPER-31120497) from Roper iPoll and load the data in R.

```{r}
#| label: load_data

poll <- read_csv("../../data/political_polling/31120497.csv")
```

2. Look at the study documentation on Roper iPoll to determine what variables are used by Ipsos to weight the respondents. Show these variables for the ten respondents with the largest weights and the ten respondents with the smallest weights. What kinds of respondents do you think Ipsos has a harder time reaching? An easier time reaching?

    *It is not necessary to evaluate every weighting variable; usually, only one or two variables determine most of the weights, even if many weighting variables are used.*

#### Largest weights

```{r}
#| label: weights_large

poll |>
  select(Weight_PID, ppgender, ppethm, ppeduc5, ppreg4, ppmsacat, ppinc7, QPID) |>
  arrange(-Weight_PID) |>
  slice_head(n = 10)
```

#### Smallest weights

```{r}
#| label: weights_small

poll |>
  select(Weight_PID, ppgender, ppethm, ppeduc5, ppreg4, ppmsacat, ppinc7, QPID) |>
  arrange(Weight_PID) |>
  slice_head(n = 10)
```

3. Create a single graph comparing the **unweighted** proportion of respondents who have a **favorable** opinion of Joe Biden (`Q1_2`) and a favorable opinion of Donald Trump (`Q1_1`). Be sure to show the margin of error on your graph.

    *Hint: you will need to use `pivot_longer()` in your answer. Review this function from our class on tidying data. In addition, it will help to rename `Q1_1` and `Q1_2` before using `pivot_longer()`.*

```{r}
#| label: fav_unweight

poll |>
  rename(Trump = Q1_1,
         Biden = Q1_2) |>
  pivot_longer(cols = c(Trump, Biden), names_to = "Candidate", values_to = "Response") |>
  count(Candidate, Response) |>
  group_by(Candidate) |>
  mutate(prop = n/sum(n),
         resp = sum(n),
         moe = 1.96 * sqrt((prop * (1 - prop))/resp)) |>
  filter(Response == "Favorable") |>
  ggplot() +
    geom_bar(aes(x = Candidate, y = prop, fill = Candidate), stat = "identity", position = "dodge") +
    geom_errorbar(aes(x = Candidate, ymin = prop-moe, ymax = prop+moe, group = Candidate), position = position_dodge(.9), width = .2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("blue", "red")) +
  labs(x = "Candidate",
       y = "Percent with Favorable Opinion",
       title = "Favorability Ratings of Biden and Trump",
       caption = "Source: August 2023 ABC News/Ipsos Poll") +
  theme_classic() +
  theme(plot.title = element_text(hjust = .5),
        legend.position = "none")
```

4. Create the same graph as in \#3, but this time using the **weighted** proportion of respondents who have a favorable opinion of Joe Biden and Donald Trump (`Q1_1`). Be sure to show the margin of error on your graph. What differences do you notice between the weighted and unweighted results? What does this tell us about which candidate's supporters are harder for Ipsos to reach?

```{r}
#| label: biden_v_trump

poll |>
  rename(Trump = Q1_1,
         Biden = Q1_2) |>
  pivot_longer(cols = c(Trump, Biden), names_to = "Candidate", values_to = "Response") |>
  count(Candidate, Response, wt = Weight_PID) |>
  group_by(Candidate) |>
  mutate(prop = n/sum(n),
         resp = sum(n),
         moe = 1.96 * sqrt((prop * (1 - prop))/resp)) |>
  filter(Response == "Favorable") |>
  ggplot() +
    geom_bar(aes(x = Candidate, y = prop, fill = Candidate), stat = "identity", position = "dodge") +
    geom_errorbar(aes(x = Candidate, ymin = prop-moe, ymax = prop+moe, group = Candidate), position = position_dodge(.9), width = .2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("blue", "red")) +
  labs(x = "Candidate",
       y = "Percent with Favorable Opinion",
       title = "Favorability Ratings of Biden and Trump",
       caption = "Source: August 2023 ABC News/Ipsos Poll") +
  theme_classic() +
  theme(plot.title = element_text(hjust = .5),
        legend.position = "none")
```

**BONUS (3 points)**: Create the same graph as in \#4, but use facets to compare favorability ratings of Biden and Trump among Democratic, Republican, and Independent voters (party ID is `QPID`).

*This question goes beyond our work in class. Think about what it means to add an additional group to our analysis. How will this affect our `count()` and `group_by()` functions?*

```{r}
#| label: pid

poll |>
  rename(Trump = Q1_1,
         Biden = Q1_2) |>
  pivot_longer(cols = c(Trump, Biden), names_to = "Candidate", values_to = "Response") |>
  filter(QPID %in% c("A Democrat", "An Independent", "A Republican")) |>
  count(QPID, Candidate, Response, wt = Weight_PID) |>
  group_by(QPID, Candidate) |>
  mutate(prop = n/sum(n),
         resp = sum(n),
         moe = 1.96 * sqrt((prop * (1 - prop))/resp)) |>
  filter(Response == "Favorable") |>
  ggplot() +
    geom_bar(aes(x = Candidate, y = prop, fill = Candidate), stat = "identity", position = "dodge") +
    geom_errorbar(aes(x = Candidate, ymin = prop-moe, ymax = prop+moe, group = Candidate), position = position_dodge(.9), width = .2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("blue", "red")) +
  labs(x = "Candidate",
       y = "Percent with Favorable Opinion",
       title = "Favorability Ratings of Biden and Trump",
       caption = "Source: August 2023 ABC News/Ipsos Poll") +
  theme_classic() +
  theme(plot.title = element_text(hjust = .5),
        legend.position = "none") +
  facet_wrap(~ QPID)
```
