---
title: "Assignment #5"
subtitle: "Fall 2023"
author: "Prof. Bell"
# format:
#   html:
#     code-tools: true
#     embed-resources: true
format: gfm
execute:
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

For this homework, we will use hypothesis tests and regression to explore Americans' views of free expression on college campuses. You will need to download the data for the [Ipsos/Knight Foundation Survey: Free Expression in America Post-2020](https://doi.org/10.25940/ROPER-31119146) from Roper iPoll.

You must submit your homework as a rendered Quarto document (HTML preferred, PDF and Word acceptable). Please ensure that all code used to generate the document (including your `setup` chunk) are visible in the report. However, extraneous code that is not required for answering the questions but that appears in your Quarto document will result in a deduction (for example, `glimpse()` should not appear anywhere in your rendered report).

Please turn in your `.qmd` file, the `_files` directory, and your rendered Quarto document as a `.zip` folder on Blackboard. When you export multiple items from Posit Cloud at once, it will automatically be exported as a `.zip`.

```{r}
#| label: setup

library(tidyverse)
library(sjstats)
library(sjPlot)
poll <- read_csv("../../data/regression/31119146.csv")
```

**BEFORE YOU BEGIN**: Make the following change to your data frame in your setup chunk before proceeding to any analysis.

```{r}
#| label: modify
#| echo: true

poll <- 
  poll |>
  mutate(Main_Weights = ifelse(!is.na(Students_Weights), Students_Weights, Main_Weights))
```

1. Generate a graph that shows the weighted topline results for the question, "Should public universities and colleges allow or prohibit a person sharing political views that are offensive to some on campus?" The relevant variable is `Q11_4` and the relevant weighting variable is `Main_Weights`.

```{r}
#| label: topline

poll |>
  filter(Q11_4 != "Skipped") |>
  count(Q11_4, wt = Main_Weights) |>
  mutate(prop = n/sum(n),
         resp = sum(n),
         moe = 1.96 * sqrt((prop * (1 - prop))/resp)) |>
  ggplot() +
  geom_bar(aes(x = Q11_4, y = prop, fill = Q11_4), stat = "identity", position = "dodge") +
    geom_errorbar(aes(x = Q11_4, ymin = prop-moe, ymax = prop+moe, group = Q11_4), position = position_dodge(.9), width = .2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("seagreen", "orangered")) +
  labs(x = "Response",
       y = "Percent",
       title = "Views of Prohibiting Offensive Political Attitudes at Colleges and Universities",
       caption = "Source: Ipsos/Knight Foundation Survey: Free Expression in America Post-2020") +
  theme_classic() +
  theme(plot.title = element_text(hjust = .5),
        legend.position = "none")
```

**BONUS:** Compare the weighted topline results for college students and non-college students in the same graph. The relevant variable is `Student1`. (Hint: Non-college students are represented with the values `No` and `NA`.)
    
```{r}
#| label: topline_college

poll |>
  filter(Q11_4 != "Skipped") |>
  mutate(College = ifelse(Student1 == "No" | is.na(Student1), "Not a student", "Student")) |>
  count(College, Q11_4, wt = Main_Weights) |>
  group_by(College) |>
  mutate(prop = n/sum(n),
         resp = sum(n),
         moe = 1.96 * sqrt((prop * (1 - prop))/resp)) |>
  ggplot() +
  geom_bar(aes(x = Q11_4, y = prop, fill = College), stat = "identity", position = "dodge") +
    geom_errorbar(aes(x = Q11_4, ymin = prop-moe, ymax = prop+moe, group = College), position = position_dodge(.9), width = .2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("forestgreen", "darkblue")) +
  labs(x = "Response",
       y = "Percent",
       title = "Views of Prohibiting Offensive Political Speech\nat Colleges and Universities",
       caption = "Source: Ipsos/Knight Foundation Survey: Free Expression in America Post-2020") +
  theme_classic(base_size = 16) +
  theme(plot.title = element_text(hjust = .5),
        legend.position = "bottom")
```

2. Conduct a weighted hypothesis test showing whether the proportions of non-college students and college students who believe that colleges and universities should **allow** offensive political speech on campus are different. You must write both the $H_0$ and the $H_A$. How do you interpret the results of your hypothesis test? The relevant variable for college students is `Student1`. (Hint: Non-college students are represented with the values `No` and `NA`.)

```{r}
#| label: hypothesis_test

poll2 <-
  poll |>
  filter(Q11_4 != "Skipped") |>
  mutate(College = ifelse(Student1 == "No" | is.na(Student1), "Not a student", "Student"),
         College = factor(College, levels = c("Not a student", "Student")),
         AllowSpeech = ifelse(Q11_4 == "Allow", 1, 0))

weighted_ttest(AllowSpeech ~ College + Main_Weights,
               poll2)
```

3. Conduct a linear probability model regression of whether colleges and universities should **allow** offensive political speech on campus on student status and at least two confounders. Please explain why you chose to control for those two confounders (i.e., why do you think that variable is a confounder?)

    How do you interpret the effect of being a current college student on support for allowing offensive political speech on campus?

    Some potential confounders include:

    * Party ID: `QPID100`
    * Gender: `ppgender`
    * Household income: `ppinc7`
    * Race/Ethnicity: `ppethm`

```{r}
#| label: regression

poll3 <-
  poll2 |>
  filter(QPID100 != "Skipped")

reg1 <- lm(AllowSpeech ~ College + QPID100 + ppinc7, data = poll3)

tab_model(reg1)
```