---
title: "Final Exam"
subtitle: "Fall 2023"
author: "Prof. Bell"
# format:
#   html:
#     code-tools: true
#     embed-resources: true
format: gfm
execute:
  warning: false
  message: false
  include: true
editor_options: 
  chunk_output_type: console
---

The American National Election Study (ANES) is a federally-funded survey of the American electorate taken at each presidential and mid-term election cycle. The survey is used by academic researchers to track American public opinion over time and is known for its high-quality data. For this assignment, we will be using data from the 2022 ANES. This particular survey is considered a pilot study, and was conducted by YouGov using a non-probability sample.

You must submit your exam as a rendered Quarto document (HTML preferred, PDF and Word acceptable). Please ensure that all code used to generate the document (including your `setup` chunk) are visible in the report. However, extraneous code that is not required for answering the questions but that appears in your Quarto document will result in a deduction (for example, `glimpse()` should not appear anywhere in your rendered report).

Please turn in your `.qmd` file, the `_files` directory, and your rendered Quarto document as a `.zip` folder on Blackboard. When you export multiple items from Posit Cloud at once, it will automatically be exported as a `.zip` file.

```{r}
#| label: setup

library(tidyverse)
library(sjstats)
library(sjPlot)
```

1. Load the data and remove any rows that are missing a value in the `weight` column.

```{r}
#| label: load_data

poll <- read_csv("../../data/final-exam/anes_pilot_2022_csv_20221214.csv") |>
  filter(!is.na(weight))
```

2. We will be working with the question, "On the whole, how satisfied are you with the way democracy works in the United States?" The relevant variable is `demsatisfy` and the values in this field represent:

    - 1 - Extremely satisfied
    - 2 - Very satisfied
    - 3 - Moderately satisfied
    - 4 - Slightly satisfied
    - 5 - Not at all satisfied
    
    Modify this variable so that we can compare those who are moderately, very, or extremely satisfied to all other responses.
    
    In addition, political party is represented by the variable `pid_x`, and the values in this field represent:
    
    - 1 - Strong Democrat
    - 2 - Democrat
    - 3 - Lean Democrat
    - 4 - Independent
    - 5 - Lean Republican
    - 6 - Republican
    - 7 - Strong Republican
    
    Modify this variable so that we can compare Democrats, Republicans, and Independents.

```{r}
#| label: wrangle

poll <-
  poll |>
  mutate(satisfied = ifelse(demsatisfy %in% c(1:3), "Satisfied", "Not Satisfied"),
         party = ifelse(pid_x %in% c(1:3), "Democrat", ifelse(pid_x == 4, "Independent", "Republican")))
```

3. Generate a nicely-formatted graph that shows the weighted topline results of the democratic satisfaction question, comparing Democrats, Republicans, and Independents. 

```{r}
#| label: topline

poll |>
  filter(!is.na(party)) |>
  count(party, satisfied, wt = weight) |>
  group_by(party) |>
  mutate(prop = n/sum(n),
         resp = sum(n),
         moe = 1.96 * sqrt((prop * (1 - prop))/resp)) |>
  ggplot() +
  geom_bar(aes(x = party, y = prop, fill = satisfied), stat = "identity", position = "dodge") +
    geom_errorbar(aes(x = party, ymin = prop-moe, ymax = prop+moe, group = satisfied), position = position_dodge(.9), width = .2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("tomato", "dodgerblue")) +
  labs(x = "Response",
       y = "Percent",
       fill = "Satisfaction",
       title = "Satisfaction with American Democracy",
       caption = "Source: 2022 American National Election Study") +
  theme_classic() +
  theme(plot.title = element_text(hjust = .5))
```

4. A "feeling thermometer" question asks respondents to rate their feelings towards a person, place, or thing on a scale of 1 to 100, where 1 is very "cold" or negative and 100 is very "warm" or positive.
    
    Conduct a weighted hypothesis test showing whether **Independents** who are satisfied with American democracy feel more warmly towards President Biden (variable `ftbiden`) than those who are not satisfied. You must write both the $H_0$ and the $H_A$. How do you interpret the results of your hypothesis test?

```{r}
#| label: hypothesis_test

poll2 <-
  poll |>
  filter(party == "Independent") |>
  mutate(satisfied = factor(satisfied, levels = c("Satisfied", "Not Satisfied")))

weighted_ttest(ftbiden ~ satisfied + weight,
               poll2,
               alternative = "greater")
```

5. Conduct a linear probability model regression of satisfaction with American democracy on political party and at least two confounders. Use Independents as your reference level for political party. Please explain why you chose to control for those two confounders (i.e., why do you think that variable is a confounder?)

    How do you interpret the effect of political party on satisfaction with American democracy?

    Some potential confounders include:

    * Birth year: `birthyr_dropdown` (be sure to convert this to age)
    * Gender: `gender` (1 - Male, 2 - Female)
    * Education: `educ` (1 - No high school, 2 - High school, 3 - Some college, 4 - Associate's degree, 5 - Bachelor's, 6 - Post-grad)
    * Family income: `faminc_new` (treat this as a continuous variable)
    * Urban-rural status: `urbanicity2` (1 - Big city, 2 - Smaller city, 3 - Suburban area, 4 - Small town, 5 - Rural area)
    
    *Hint: Make sure that categorical variables are treated as such in the regression, and not as continuous variables.*

```{r}
#| label: regression

poll3 <-
  poll |>
  mutate(satisfied = ifelse(demsatisfy %in% c(1:3), 1, 0),
         educ = factor(educ),
         party = relevel(factor(party), ref = "Independent"))

reg1 <- lm(satisfied ~ party + educ + faminc_new, data = poll3)

tab_model(reg1,
          pred.labels = c("Intercept", "Democrat", "Republican", "High School", "Some College", "Associate's", "Bachelor's", "Post-grad", "Income"),
          dv.labels = "Satisfied with Democracy")
```